\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2010-11-24 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed all the warnings detected by IBM xlf

\begin{description}
\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
\end{description}
On a new system, it is definitely worth experimenting with single
versus double precision simulations to determine which is faster.
Note that on many current processors (e.g., Intel, AMD, IBM Power),
single precision calculations are often significantly faster; the
difference can typically be 10\% to 25\%. It is therefore often worth
using single precision if you can. We recommend running the same calculation
once in single precision and in double precision on your system and
then comparing the seismograms. If they are identical, you should probably
select single precision for your future runs.

When running on an SGI add ``\texttt{setenv TRAP\_FPE OFF}'' to your
.cshrc file \textit{before} compiling in order to turn underflow trapping
off.

% obsolete
%Finally, before compiling make sure that the subdirectories \texttt{obj}
%and \texttt{OUTPUT\_FILES} exist within the directory with the source
%code (\texttt{SPECFEM3D\_V2.0.0}). The \texttt{go\_mesher} script
%discussed in Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} automatically
%takes care of creating the \texttt{OUTPUT\_FILES} directory.

Note that if you run very large meshes on a relatively small number
of processors, the memory size needed on each processor might become
greater than 2 gigabytes, which is the upper limit for 32-bit addressing;
in this case, on some compilers you may need to add \char`\"{}\texttt{-mcmodel=medium}\char`\"{}
to the compiler options otherwise the compiler will display an error
message.

@@

%\noindent \begin{flushleft}
%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi}
\par\end{centering}

@@

%\par\end{flushleft}

@@

%obsolete...
%\item [{\texttt{TOPOGRAPHY}}] Set to \texttt{.true.} if topography and
%bathymetry should be incorporated based upon model ETOPO5 \citep{Etopo5}.
%This feature adds no cost to the simulation.

@@

%obsolete...
%\item [{\texttt{RECORD\_LENGTH\_IN\_MINUTES}}] Choose the desired record
%length of the synthetic seismograms (in minutes). This controls the
%length of the numerical simulation, i.e., twice the record length
%requires twice as much CPU time. This feature is not used at the time
%of meshing but is required for the solver, i.e., you may change this
%parameter after running \texttt{xgenerate\_databases}.

@@

%% Another file generated by \texttt{xgenerate\_databases} is the header file \texttt{OUTPUT\_FILES/values\_from\_mesher.h}.
%% This file specifies a number of constants and flags needed by the
%% solver. These values are passed statically to the solver for reasons
%% of speed. Some useful statistics about the mesh are also provided
%% in this file.

%% For a given model, set of nodes, and set of parameters in \texttt{Par\_file},
%% one only needs to run the mesher once and for all, even if one wants
%% to run several simulations with different sources and/or receivers
%% (the source and receiver information is used in the solver only).

@@

%obsolete
%For reasons of speed, the solver uses static memory allocation.
%Therefore it needs to be recompiled (type `\texttt{make clean}' and
%`\texttt{make specfem3D}') every time one reruns \texttt{xgenerate\_databases}. To compile
%the solver one needs a file created by \texttt{xgenerate\_databases} in the directory
%\texttt{OUTPUT\_FILES} called \texttt{values\_from\_mesher.h}, which
%contains parameters describing the static size of the arrays as well
%as the setting of certain flags.

@@

\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
%\item the attenuation control parameters \texttt{ATTENUATION} and \texttt{USE\_OLSEN\_ATTENUATION}
\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS}
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
\texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.

@@

%<YANGL
\chapter{Noise Cross-correlation Simulations}

@@

Even if this script does not work on your workstation, the procedure it describes is universal.
You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
which should contain enough explanations for all the commands.
%>YANGL

@@

% uses obsolete routines...
%\item [\texttt{xcreate\_movie\_GMT}]
%From \texttt{create\_movie\_GMT.f90},
%outputs ascii xyz files, convenient for use with GMT. This code uses
%significantly less memory than \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}
%and is therefore useful for high resolution runs.

@@

\begin{itemize}
\item The processors/nodes to be used for each run are assigned dynamically
by the scheduler, based on availability. Therefore, in order for the
\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
have access to the same database files (if they are stored on hard
drives local to the nodes on which the code is run), they must be
launched in sequence as a single job.
\item On some systems, the nodes to which running jobs are assigned are
not configured for compilation. It may therefore be necessary to pre-compile
both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
%obsolete
%A small program provided in the distribution
%called \texttt{\small create\_header\_file.f90} can be used to directly
%create \texttt{\small in\_out\_files/OUTPUT\_FILES/values\_from\_mesher.h} using
%the information in the \texttt{\small in\_data\_files/Par\_file} without having
%to run the mesher (type \texttt{\small `make}{\small{} }\texttt{\small create\_header\_}~\\
%\texttt{\small file}' to compile it and `\texttt{\small xcreate\_header\_file}'
%to run it; refer to the sample scripts below). The solver can now
%be compiled as explained above.
\item One feature of schedulers/queuing systems is that they allow submission
of multiple jobs in a ``launch and forget'' mode. In order to take
advantage of this property, care needs to be taken that output and
intermediate files from separate jobs do not overwrite each other,
or otherwise interfere with other running jobs.
\end{itemize}
Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
directory and can straightforwardly be modified and adapted to meet
more specific running needs.

@@

\begin{lyxcode}
%{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}
{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
{\footnotesize \par}
\end{lyxcode}
%which has cut all the sac files between 0 and 180 seconds, filtered
%them between 2 and 40 seconds, transfered them into displacement records
%using the polezero files in \texttt{dir} directory, picked the first
%P and S arrivals, and added suffix `\texttt{bp}' to the file names.

@@

\begin{lyxcode}
%{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}
{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
{\footnotesize \par}
\end{lyxcode}
%which will convert the synthetics into SAC format, add event and station
%information into the SAC headers, cut the SAC files between 0 and
%180 seconds, filter them between 2 and 40 seconds, pick the first
%P and S arrivals, and add the suffix `\texttt{bp}' to the file names.
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
to the file names.

>>>>

\begin{description}
\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
\end{description}
On a new system, it is definitely worth experimenting with single
versus double precision simulations to determine which is faster.
Note that on many current processors (e.g., Intel, AMD, IBM Power),
single precision calculations are often significantly faster; the
difference can typically be 10\% to 25\%. It is therefore often worth
using single precision if you can. We recommend running the same calculation
once in single precision and in double precision on your system and
then comparing the seismograms. If they are identical, you should probably
select single precision for your future runs.\\

When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
with the following options:\\

\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-path=...}.\\

On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
in order to turn underflow trapping off.\\

Note that if you run very large meshes on a relatively small number
of processors, the memory size needed on each processor might become
greater than 2 gigabytes, which is the upper limit for 32-bit addressing;
in this case, on some compilers you may need to add \char`\"{}\texttt{-mcmodel=medium}\char`\"{}
to the compiler options otherwise the compiler will display an error message; on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.

@@

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi}
\par\end{centering}

@@

@@

@@

@@

@@

@@

\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS}
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
\texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.

@@

\chapter{Noise Cross-correlation Simulations}

@@

Even if this script does not work on your workstation, the procedure it describes is universal.
You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
which should contain enough explanations for all the commands.

@@

@@

\begin{itemize}
\item The processors/nodes to be used for each run are assigned dynamically
by the scheduler, based on availability. Therefore, in order for the
\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
have access to the same database files (if they are stored on hard
drives local to the nodes on which the code is run), they must be
launched in sequence as a single job.
\item On some systems, the nodes to which running jobs are assigned are
not configured for compilation. It may therefore be necessary to pre-compile
both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
\item One feature of schedulers/queuing systems is that they allow submission
of multiple jobs in a ``launch and forget'' mode. In order to take
advantage of this property, care needs to be taken that output and
intermediate files from separate jobs do not overwrite each other,
or otherwise interfere with other running jobs.
\end{itemize}
Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
directory and can straightforwardly be modified and adapted to meet
more specific running needs.

@@

\begin{lyxcode}
{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
{\footnotesize \par}
\end{lyxcode}

@@

\begin{lyxcode}
{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
{\footnotesize \par}
\end{lyxcode}
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
to the file names.
